### 2014-03-02 paper/retrofit.tex
Fix many notational issues and other things in retrofit.

Our overall plan will be to take a \textbf{{\color{red} target language}} with a formal
operational semantics and combine it with an \textit{{\color{blue} information
flow control language}}.  For concreteness, we use a simple, untyped,
ML-like language with mutable references as our target language.
However, our technique will be indifferent to the actual target language
chosen: it need only be prepared for interoperability by being expressed
in the style of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220}.  This procedure will
equip the target language with primitives for implementing dynamic
information control in the style of LIO~\cite{lio}, which we will
describe in more detail shortly.

We have typeset nonterminals of the target language using \textbf{{\color{red}
bold font}} while the nonterminals of the IFC language have been typeset
with \textit{{\color{blue} italic font}}.  Readers are encouraged to view
a color copy of this paper, where target language nonterminals are colored red
while IFC language nonterminals are colored blue.

@@

In Figure~\ref{fig:ml}, we give a simple, untyped ML-like language with
mutable references and errors, prepared for combination with an
information flow control language.  The presentation is standard, but it
is worth paying some attention to the Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} used to define the
operational semantics of the system.  Our language defines an evaluation
context |tE|; however, the evaluation rules have been expressed in terms
of a different evaluation context $\mathcal{E}_{\tar{\Sigma}}$.  To
derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, tE[ te ]|.
However, when we combine this language with an IFC language, we must
reinterpret the meaning of this evaluation context.
In general, we require that a target language be expressed in terms
of some global machine state |tS|, some evaluation context |tE|,
some expressions |te|, some set of values |tv| and a reduction
relations on full configurations $|tS| \times |tE| \times |te|$.

@@

\and
\inferrule[T-app]
{ } {| conf tS ((\x.te) tv) -> conf tS ({tv/x} te) |}

@@

\and
\inferrule[T-deref]
{ |(ta, tv) in tS| }
{| conf tS (!ta) -> conf tS (tv) |}

@@

Information flow control seeks to enforce a security policy known as
\emph{non-interference}: public results must not depend on secret inputs.
Modern, dynamic IFC systems usually achieve this by associating a \emph{label}
with data, which encodes a policy of whether or not data tagged with one
label can flow to another label (written as |canFlowTo|).
Since our embedding is agnostic
to the choice of labeling scheme, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

In Figure~\ref{fig:ifc}, we give the syntax and \emph{single-threaded}
evaluation rules for a minimal information-flow control language.  The
evaluation context $\mathcal{E}_{\ifc{\Sigma}}$ represents the context
of the current task which is running, including on the right-hand side
the current label $l$ (which protects the program counter of the task)
and the task ID $i$.

@@

The form of \textbf{send}/\textbf{recv} are \Red{inspired by methods of inter-worker communication
in web browser APIs} \Red{cite}, however in Section~\ref{sec:extensions} we
discuss other possible choices for information flow control aware communication
primitives.  Their rules are different from non-IFC systems
and worth commenting on: when a task places
a message on the message queue of receiving task, it also labels that
message with a label |il'| (which must be above the tasks current label |il|)
When the recipient reads out messages from its queue,
messages whose labels do not
flow to |il| are filtered out (denoted by |filter il Q|).
Specifically, if |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more secret
than their current label would allow.

@@

\begin{figure}
\begin{code}
lop  ::=  canFlowTo | lub | glb
iv   ::=  id | il | true | false | unit
ie   ::=  iv | ix | ie lop ie | getLabel | setLabel ie
       |  sandbox ie | send ie ie ie | ix1, ix2 <= recv. ie ie
iE   ::=  iHole | iE lop ie | iv lop iE | setLabel iE | sandbox iE 
       |  send iE ie ie | send iv iE ie | send iv iv iE
       |  ix1, ix2 <= recv. ie ie
io   ::=  (il, id ie)
Q    ::=  nil | io , Q
iS   ::=  emptyset | bracket iS (mapsto id Q)
\end{code}

@@

@@

\and
\inferrule[I-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = bracket iS (mapsto id' (il', id,  ie) , Q)|
}
{|
niconf iS  id il (send id' il' ie)
->
niconf iS' id il unit
|}

\and
\inferrule[I-recv]
{
| filter il (iS(id)) = io1 , ... , iok , (il', id', ie')|\\
| iS' = bracket iS (mapsto id (io1 , ... , iok))|\\
}
{|
niconf iS  id il (ix1, ix2 <= recv. ie1 ie2)
->
niconf iS' id il ({ie'/ix1,id'/ix2} ie1)
|}

\and
\inferrule[I-noRecv]
{
| filter il (iS(id)) = nil|\\
| iS' = bracket iS (mapsto id nil)|
}
{|
niconf iS  id il (ix1, ix2 <= recv. ie1 ie2)
->
niconf iS' id il ie2
|}

@@

\Red{TODO[ds]:
  - move the async recv to text and out of the figure
  - clarify discussion on exceptions
  - fix the toLabled
  - add taskIDs to ruls
}
Finally, in Figure~\ref{fig:embedding},
we provide all of the rules responsible for actually carrying out the embedding.
The most important feature of this embedding is that every task maintains its own
copy of the target language global state and evaluation context, thus
enforcing isolation between various tasks.  In more detail:

@@

\begin{mathpar}
\inferrule[I-sandbox]
{ 
|iS' = bracket iS (mapsto id' nil)|\\
|tS' = klone(tS)|\\
|it1 = fullconf id il tS (iniE id')|\\
|itnew = fullconf id' il tS' (TI ie)|\\
|ap fresh id'|
}
{|
iconf iS (fullconf id il tS (iniEi (sandbox ie)), ldots)
.->
iS'; sched F (it1, ldots, itnew)
|}

\and
\inferrule[I-done]
{|
|}
{|
iconf iS (fullconf id il tS iv, ldots)
.->
iS ; sched D (fullconf id il tS iv, ldots)
|}

\and
\inferrule[I-noStep]
{|
iconf iS (it, ldots) .-/> 
|}
{|
iconf iS (it, ldots)
.->
iS ; sched N (it, ldots)
|}

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of both languages to allow for embeddings terms in one language to be embedded in the other.
    \item We reinterpret |CE| to be evaluation contexts on task pools, providing definitions for $\mathcal{E}_{\tar{\Sigma}}$ and $\mathcal{E}_{\ifc{\Sigma}}$.  These rules only operate on the first task in the task pool, which by convention, is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task pools.  The correspondence is simple: a task executes a step and then is rescheduled in the task pool according to schedule policy $\alpha$.  Not all schedulers preserve non-interference, so we provide two specific schedules in Figure~\ref{fig:scheduler}.
    \item Finally, we define some rules for scheduling, handling sandboxing tasks (which interact with the state of the target language) and intermediating between the borders of the two languages.
\end{itemize}

The \textsc{I-sandbox} rule is of particular interest, as it is intended to allow users
to create new tasks.  When a new task is created, there is a question of what the
global state of the new task should be; this is defined in terms of a function $\kappa$.
Conservatively, $\kappa$ may be simply thought of as the identity function, in which
case the semantics of sandbox are such that the state of the target language is \emph{cloned}
when sandboxing occurs.  However, this is not necessary: it is also valid for $\kappa$
to remove entries from the state.  In Section~\ref{sec:concrete}, we give a more detailed
discussion of the implications of the choice of $\kappa$.

>>>>

Our overall plan will be to take a \textbf{{\color{red} target language}} with a formal
operational semantics and combine it with an \textit{{\color{blue} information
flow control language}}.  For concreteness, we use a simple, untyped,
ML-like language with mutable references as our target language.
However, our technique will be indifferent to the actual target language
chosen: it need only be prepared for interoperability by being expressed
in the style of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220},
and for simplicity, we require the semantics to be deterministic.
This procedure will
equip the target language with primitives for implementing dynamic
information control in the style of LIO~\cite{lio}, which we will
describe in more detail shortly.

We have typeset nonterminals of the target language using \textbf{{\color{red}
bold font}} while the nonterminals of the IFC language have been typeset
with \textit{{\color{blue} italic font}}.  Readers are encouraged to view
a color copy of this paper, where target language nonterminals are colored \textbf{{\color{red} red}}
while IFC language nonterminals are colored \textit{{\color{blue} blue}}.

@@

In Figure~\ref{fig:ml}, we give a simple, untyped ML-like language with
mutable references, prepared for combination with an
information flow control language.  The presentation is standard, but it
is worth paying some attention to the Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} used to define the
operational semantics of the system.  Our language defines an evaluation
context |tE|; however, the evaluation rules have been expressed in terms
of a different evaluation context |curlyRedE|,
following the approach of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220} to combine operational
semantics of multiple languages.  To
derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, bracket tE te|.
However, when we combine this language with an IFC language, we must
reinterpret the meaning of this evaluation context.
In general, we require that a target language be expressed in terms
of some global machine state |tS|, some evaluation context |tE|,
some expressions |te|, some set of values |tv| and a deterministic
reduction relation on full configurations $|tS| \times |tE| \times |te|$.

@@

\inferrule[T-app]
{ } {| conf tS ((\x.te) tv) -> conf tS ({tv/x} te) |}

@@

\and
\inferrule[T-deref]
{ |memberf (ta, tv) tS| }
{| conf tS (!ta) -> conf tS (tv) |}

@@

Information flow control seeks to enforce a security policy known as
\emph{non-interference}: public results must not depend on secret inputs.
Modern, dynamic IFC systems usually achieve this by associating a \emph{label}
with data, which encodes a policy of whether or not data tagged with one
label |l1| can flow to another label |l2| (written as |canFlowTo2 l1 l2|).
Our embedding is agnostic to the choice of labeling scheme; we only
require the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056}
with the partial order |canFlowTo|.
In this paper, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

In Figure~\ref{fig:ifc}, we give the syntax and \emph{single-threaded}
evaluation rules for a minimal information-flow control language.  The
evaluation context |curlyBlueE| represents the context
of the current task which is running, including on the right-hand side
the current label |il| (which protects the program counter of the task)
and the task ID |id|.

@@

We consider asynchronous communication primitives that allow different
tasks to exchange messages.
When a task is sending a message using |send|, it also labels that
message with a label |il'| (which must be above the tasks current label |il|).
Now, messages can only be received by a task if its current label is
at least as high as the label of the message.

Specifically, receiving a message using
\[ |recvf ix1 ix2 ie1 ie2| \]
binds the senders task identifier and the message
to local variables |ix1| and |ix2|, respectively, and then executes |ie1|.
Otherwise the message gets removed from the queue by rule \textsc{I-noRecv},
and that tasks continues its execution with |ie2|.
We denote the filtering of the message queue by |filter il Q|,
which is defined as follows.
If |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more secret
than their current label would allow.

The particular choice of communication primitives has been motivated by
inter-worker communication in web browser APIs~\Red{cite}, but
other choices such as blocking communication or labeled channels
would also be possible.

@@

\begin{figure}
\begin{code}
lop  ::=  canFlowTo | lub | glb
iv   ::=  id | il | true | false | unit
ie   ::=  iv | ix | ie lop ie | getLabel | setLabel ie | getTaskId
       |  sandbox ie | send ie ie ie | recvf ix ix ie ie
iE   ::=  iHole | iE lop ie | iv lop iE | setLabel iE | sandbox iE
       |  send iE ie ie | send iv iE ie | send iv iv iE
io   ::=  (il, id ie)
Q    ::=  nil | io , Q
iS   ::=  emptyset | bracket iS (mapsto id Q)
\end{code}

@@

\inferrule[I-getTaskId]
{ }
{|
niconf iS id il getTaskId
->
niconf iS id il id
|}

@@

\and
\inferrule[I-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = bracket iS (mapsto id' (il', id,  iv) , Q)|
}
{|
niconf iS  id il (send id' il' iv)
->
niconf iS' id il unit
|}

\and
\inferrule[I-recv]
{
| (filter il (iS(id))) = io1 , ... , iok , (il', id', iv)|\\
| iS' = bracket iS (mapsto id (io1 , ... , iok))|\\
}
{|
niconf iS  id il (recvf ix1 ix2 ie1 ie2)
->
niconf iS' id il ({iv/ix1,id'/ix2} ie1)
|}

\and
\inferrule[I-noRecv]
{
| filter il (iS(id)) = nil|\\
| iS' = bracket iS (mapsto id nil)|
}
{|
niconf iS  id il (recvf ix1 ix2 ie1 ie2)
->
niconf iS' id il ie2
|}

@@

\Red{TODO[ds]:
  - move the async recv to text and out of the figure
  - clarify discussion on exceptions
  - fix the toLabled
}
Finally, in Figure~\ref{fig:embedding},
we provide all of the rules responsible for actually carrying out the embedding.
The most important feature of this embedding is that every task maintains its own
copy of the target language global state and evaluation context, thus
enforcing isolation between various tasks.  In more detail:

@@

\begin{mathpar}
\inferrule[I-sandbox]
{ 
|iS' = bracket iS (mapsto id' nil)|\\
|tS' = klone(tS)|\\
|it1 = fullconf id il tS (iniE id')|\\
|itnew = fullconf id' il tS' (TI ie)|\\
|ap fresh id'|
}
{|
iconf iS (fullconf id il tS (iniEi (sandbox ie)), ldots)
.->
iS'; sched box (it1, ldots, itnew)
|}

\and
\inferrule[I-done]
{|
|}
{|
iconf iS (fullconf id il tS iv, ldots)
.->
iS ; sched done (fullconf id il tS iv, ldots)
|}

\and
\inferrule[I-noStep]
{|
iconf iS (it, ldots) .-/> 
|}
{|
iconf iS (it, ldots)
.->
iS ; sched nostep (it, ldots)
|}

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of both languages to allow for embeddings terms in one language to be embedded in the other.
    \item We reinterpret |CE| to be evaluation contexts on task pools, providing definitions for |curlyRedE| and |curlyBlueE|.  These rules only operate on the first task in the task pool, which by convention is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task pools.  The correspondence is simple: a task executes a step and then is rescheduled in the task pool according to schedule policy |alpha|.  Not all schedulers preserve non-interference, and we provide two specific schedules in Figure~\ref{fig:scheduler}.
    \item Finally, we define some rules for scheduling, handling sandboxing tasks (which interact with the state of the target language),
    primitives for asynchronously communicating between
    tasks and intermediating between the borders of the two languages.
\end{itemize}

The \textsc{I-sandbox} rule allows to create a new isolated task that
executes separately from the existing tasks (and can be communicated
with via |send| and |recv|).  When the new task is created, there
is the question of what the target language state of the new task should
be.  Our rule is stated generically in terms of a function |klone|.
Conservatively, |klone| may be simply thought of as the identity
function, in which
case the semantics of sandbox are such that the state of the target language is \emph{cloned}
when sandboxing occurs.  However, this is not necessary: it is also valid for |klone|
to remove entries from the state.  In Section~\ref{sec:concrete}, we give a more detailed
discussion of the implications of the choice of |klone|, but all our
non-interference claims will hold regardless of the choice of |klone|.
